\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-05-03 trunk/arch.tex
updates

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while maintaining its reactivity.

Traditionally planning has been deemed computationally expensive with
planner performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for. In such a situation, the agent may thrash
\footnote{\kcomment{In such an event as the planner works towards goal
    satisfaction, the initial conditions leading to achieve that goal
    are no longer valid. The planner then tries to achieve a different
    goal which too has to be discarded similarly. And so on.}}  if the
internal state of the plan gets out of synch with the actual state of
the world.

\kcomment{A important consequence of this has been that} the planning
problem managed {\em in situ} has traditionally remained fairly
detached from low-level control. Most adaptations at the lower-level
reactive layers in contrast are managed by specialized components (for
example \texttt{LAAS architecture} \cite{alami:1998p820,Ingrand07} and
\texttt{CLARATy} \cite{Nesnas:2003do}) that rely on different
representations for modeling local control behavior or worse do not
have any explicit formal agent behavior model. The consequent
different techniques for specifying each layer in the architecture
results in duplication of effort and a diffusion of knowledge leading
to significant design and integration issues \cite{DS1report}.
\texttt{IDEA} mitigated most of these issues by tightly integrating
planning and execution in a single representational and computational
framework with \kcomment{an} unified declarative model. Our approach
with \rx builds on this design methodology albeit in a substantially
systematic manner. The \rx architecture contributes to agent
architectures over and beyond \texttt{IDEA} in the following
\kcomment{significant} ways:

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops, called {\em reactors}. Each reactor has its own
  scope both functionally (one for example focusing on the conversion
  of high level waypoints into low-level commands) and temporally as
  each reactor declares its planning look-ahead $\pi$ and expected
  latency $\lambda$ before being able to produce \kcomment{a} plan to
  the agent depending on where it is located in a reactor dependency
  graph such as shown in Fig. \ref{fig:agent}. A reactor closer to the
  hardware in the hierarchy, is expected to deliberate less and be
  highly responsive to observations from the world. This implies lower
  values of $\pi$ and $\lambda$ (see Section \ref{sec:arch:trex}
  below).

\item \emph{Coupling} a tighter integration between the planning
  process and world evolution to ensure that planning occurs
  \emph{during continuous} state update in the {\em same}
  \kcomment{plan} database used for planning. By doing so, the planner
  is better informed in its deliberation despite a larger latency than
  the changing world.
\end{itemize}

Systematic partitioned inference is a major contribution of
\rxe. However that is not the focus of this chapter with details
covered in \cite{py10, rajan12}. Coupled interaction between planning
and the world on the other hand, has a significant implication on how
planning can be integrated in such frameworks. In this chapter we show
how using the \eu planner can be coupled to our framework by first
linking concepts in temporal planning with the \rx framework. Note
that in the following treatise, we call the formal semantic framework
of \rx as the 'agent'. Reactors constitute an entity within this agent
and are standalone blocks of inference. The scope, detail and number
of reactors are application specific and outside the scope of this
chapter.

@@

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each such
state variable. In this framework each reactor is seen as a 'black
box' providing its own {\em internal} state variable which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado AUV, we use a
tick of $1$ second. In order to identify its state a reactor can
subscribe to {\em external} state variables owned by other reactors,
for which it will receive new updates when the owner reactor generates
its {\em internal} state.

@@

As the execution frontier, $\tau$ advances, the \rx agent propagates
the current state information across all reactor state variables
through the {\em synchronization} process.  The sequencing of the
reactors synchronization is identified by traversing a reactor
dependency graph from the least dependent -- ones with no {\em
  external} state variables -- to the most, \ie which typically do not have
{\em internal} state variables. Fig. \ref{fig:agent} shows an example
of such a dependency graph implemented as a directed acyclic graph.

Conversely each reactor can request a future change on one of its {\em
  external} state variables which will be transformed as an {\em
  internal} goal to the owner reactor(s).  Such \emph{goal
  dispatching} is expected to produce a deliberation phase for the
receiving reactor which can cyclically in turn, produce a plan
impacting its {\em external} state variables' future that will
propagate down following the same mechanism.

@@

\item $\pi$ is the planning \emph{look-ahead} of the reactor. which is
  indicative of the number of ticks in the future this reactor is
  looking ahead during deliberation.

@@

\begin{definition}
  The {\em execution latency} of a {\em state variable} $s$ which is
  internal to the reactor $r$ is the maximum expected delay before a
  goal on this state variable can be included in its plan and executed
  by the reactor $r$. It is defined recursively as: 
  \begin{equation*}
    \lambda_{exec}(s) = \lambda(r) + \max_{e \in exts(r)}
    \lambda_{exec}(e) 
  \end{equation*}
  Where  $exts(r)$ is the set of the state variables which are
  {\em externals} to $r$. 
\end{definition}

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on {\em external} state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable.

@@

This window is used by \rx to identify when to dispatch a request to a
{\em reactor} as a new goal. This happens only if the request start
time -- represented as the interval of possible start times --
intersects this window. Requests further in the future are buffered by
the agent until this time window advances for an overlap. This
provides an elegant mechanism to request a change from the {\em
  external} state variable of a reactor while avoiding flooding the
owner reactor of this state variable.

The choice of both the latency and look-ahead of a reactor can be
delicate as one is dependent on the other. Needless to say, latency is
highly dependent on exogenous paramaters such as the host computers
computing power and the number of reactors needing to deliberate at
once, among others. As the design stands, these variables should be
evaluated carefully by the system designer balancing agent efficacy.

@@

\item The future objectives a reactor has on \textsf{Internal} state
  will likely imply sub-objectives to be reported to the owner of its
  {\em external} state variables depending on their current
  values. Should the \textsf{Pilot} want to visit a specific
  \textsf{Waypoint} -- given its current position as provided by
  \textsf{Sensor\_Data}, it can identify a sequence of
  \textsf{Nav\_cmd} states that should eventually help it reach this
  location. %\comment{this explanation could do with a figure which I
    %believe we have used previously}

@@

Given the above example, \rx abstracts reactor interaction by
constraining them to state variables. Each reactor is therefore
agnostic where its {\em external} state information is managed as long
as this information is available and properly maintained for managing
{\em internal} state information. This {\em internal} state in turn
may be used by other reactors. Such a formalism also constrains
reactor interaction to state information exchange which comes in two
forms as illustrated in Fig. \ref{fig:trex:flow}:

@@

\item {\em Observation on current state values}: This information
  propagates \emph{up} in the reactor dependency graph at the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variables for this tick so they can
  compute their {\em internal} state variable values for this same
  tick. By doing so we propagate a consistent view of that state of
  the world throughout all the reactors as time
  advances. \kcomment{Observations can therefore be both sensor
    measurements as well state passed up from other reactors. The
    strict rules of timeline ownership enable a clear policy for
    conflict resolution: observations dominate expectations.}

\item {\em Goal request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy own
  {\em internal} objectives, a reactor may have a plan that relies on
  future values of one of its {\em external} state. When such a plan
  is identified the agent ensures that this part of the plan is
  provided as a goal to the owner of the given {\em external} state
  variable(s).

@@

The overall execution cycle of the agent is also abstracted out as a
continuous planning/deliberation cycle between all reactors. At the
agent interface level, each reactor provides only two abstract calls
that conceptualize execution:

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example, one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  inconsistent. Backtracking is included in the step should it be such
  a case.  The returned value by the \texttt{step} function is a
  boolean which indicates whether the reactor needs an extra step to
  deliberate or if this last step resulted in a complete plan within
  the reactor's look-ahead. The part of the plan that covers the
  reactor's {\em external} timelines is provided to be dispatched as a
  goal.

@@

Both these functions have different scope and temporal constraints.
The \texttt{synchronize} call is considered as atomic and will focus
only on state inference for the current tick by identifying the
current value of {\em internal} state variables from the latest
updates on {\em external} state variables. This process propagates up
in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process focused on
the future evolution of state variables; it can take multiple steps
spanning multiple ticks for a reactor to produce a plan that is both
complete and valid. How the reactor identifies that its plan is
complete and valid is its responsibility and depends on the way each
reactor resolves its plan at each {\em step}. The general idea is that
the reactor is able to find a valid strategy in order to satisfy as
many of its {\em internal} goals as possible. For \eue, this means
that the solver is unable to identify any further flaws to be resolved
in the plan.

When such plan is identified, its {\em external} timeline(s) are then
dispatched to the corresponding reactors which own these times and
which in turn can start their own deliberation \texttt{step}s to
satisfy these new objectives. Consequently execution occurs as these
goals and plans propagate down in the reactor hierarchy until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal into a command for
execution in hardware.

\label{sec:delib-exec-as}
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \rx reactor execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

@@

\subsubsection{Handling Failure in \rx}
\label{sec:rx-reactor-failure}

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors are considered \emph{rejectable} by the
reactor that manages them. As a result in such a situation, the
failure to find a plan for these goals, is \emph{not} considered a
critical failure of the reactor. It will only result \kcomment{in}
these goals being ignored until newer goals are received.

Take for example our AUV agent from Fig. \ref{fig:agent}, and assume
that the \texttt{Sampling} reactor failed to synchronize. In such a
case this reactor is destroyed and both \texttt{Sampling\_policy} and
\texttt{Relevant\_data} timelines generate the \texttt{Failed}
observation as illustrated in Fig. \ref{fig:fail0}.  The
\texttt{Survey controller} will receive this \texttt{Failed}
observation on its {\em external} \texttt{Sampling\_policy} timeline
which indicates that this state variable is no longer managed by its
owner reactor. This leads to two potential outcomes based on how the
reactor is implemented:

@@

\item The \texttt{Survey controller} model allows it to integrate the
  \texttt{Failed} observation. For example it can identify that,
  subsequent use of the water sampler is no longer possible.  The \rx
  agent can continue to navigate since \texttt{Pilot} is still
  available. In such a case the agent structure remains as in
  Fig. \ref{fig:fail0} with no consequent impact.

@@

\begin{figure}[b]
  \centering
  \subfloat[\small When the \texttt{Sampling} reactor fails to
  synchronize, it is removed from the reactor dependency graph along
  with all its {\em internal} timelines while generating a \texttt{Failed} observation]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.pdf}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent \kcomment{from} Fig. \ref{fig:agent}}
\end{figure}

@@

\begin{figure}[!b]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for. Arrows indicate temporal
      constraints. }
  \label{fig:shop:exec}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs the reactor of this observation which is
added to the partial plan as a fact -- starting exactly at $\tau'$ and
lasting for an unknown duration which is at least one tick. As this
new observation is integrated in the plan through synchronization, the
solver can identify the similarity between this observation and the
next token in the existing plan. Instead of initiating an exhaustive
search of all possible implications, the synchronization process
starts its execution with the assumption that these new observations
are consistent with the currently maintained plan. As a consequence,
at the end of synchronization the plan depicted in Fig.
\ref{fig:shop:exec1} is likely. This is a result of propagation of
information provided by the new observation in the existing plan while
restricting the start time of \texttt{Going} and propagating temporal
constraints, making \texttt{Go} to start {\em exactly} at $\tau'$. In
doing so, the \eu solver's focus on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}) takes advantage of the existing plan to
evaluate a solution that assumes that the observation is consistent
with what was decided by the reactor.

% \begin{figure}[!htb]
%   \centering
%   \includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t1}
%   \caption{Shopping example : Result of synchronization after the {\em
%       Going} observation was received by the reactor with the plan
%     showed in Fig. \ref{fig:shop:exec0}.}
%   \label{fig:shop:exec1}
% \end{figure}

This choice is directed not only by heuristic choices made in the
solver but also enforced by the heuristic order of the decision point
resolution choices for the {\em current state flaw} shown in
Proposition \ref{prop:csf:resolve}. As noted, the first two choices
are biased towards a conservative view with respect to the existing
plan. The heuristics first attempt to extend the previous observation
for this {\em internal} timeline to avoid early commitment. If this
fails \rx attempts to start the next existing token in the plan to
advance plan execution. Both choices attempt to resolve state with the
least amount of plan perturbation by constraining further the start
and end time of existing tokens generated by previous deliberations
steps. By doing so, in a nominal situation, synchronization remains
limited to restrict time points in the plan and propagates their
impact. Most of the computational effort of creating new tokens in the
plan and scheduling them onto timelines are managed during
deliberation steps. Consequently, most of the time synchronization is
limited to very few nodes in the \eu search space often requiring no
backtracking. Fig. \ref{fig:shop:exec} shows how the existing plan can
support the synchronization process. The observation \texttt{Going}
occurring at $\tau'$ is fully compatible with the token already present
in our plan. By updating the plan and propagating equality constraints,
the resolution of the {\em current state flaw} for the {\em Agent}
timeline becomes obvious as a solution, as the token \texttt{Go} from
the plan presents itself.

@@

\begin{figure}[!b]
  \centering
  \subfloat[Conflicting Observation]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[Relaxed Partial Plan]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}

>>>>

A prime motivation behind the \rx architecture was to design a
controller which could bring automated planning closer to low-level
control of the vehicle while \kcomment{ensuring}
reactivity. Traditionally planning has been deemed computationally
expensive with planner performance restricting robot reactivity
\cite{ghallab04,Dias:2003ua}. In control architectures where task
planning is embedded, planning consequently remains at an abstract
level with the assumption that it would otherwise impede system
reactiveness and that the environment can change at a faster rate than
the planner can plan for. In such a situation, the agent may thrash
\footnote{\kcomment{In such an event even as the planner works towards
    goal satisfaction, the initial conditions leading to achieve that
    goal are no longer valid. The planner then tries to achieve a
    different goal which too has to be discarded similarly. And so
    on.}}  if the internal state of the plan gets out of sync with the
actual state of the world.

\kcomment{An important consequence of this has been that} the planning
problem managed {\em in situ} has traditionally remained fairly
detached from low-level control. Most adaptations at the lower-level
reactive layers in contrast are managed by specialized components (for
example \texttt{LAAS architecture} \cite{alami:1998p820,Ingrand07} and
\texttt{CLARATy} \cite{Nesnas:2003do}) that rely on different
representations for modeling local control behavior or worse do not
have any explicit formal agent behavior model. The consequent
different techniques for specifying each layer in the architecture
results in duplication of effort and a diffusion of knowledge leading
to significant design and integration issues \cite{DS1report}.
\texttt{IDEA} mitigated most of these issues by tightly integrating
planning and execution in a single representational and computational
framework with \kcomment{an} unified declarative model. Our approach
builds on this design methodology albeit in a substantially systematic
manner.

The \rx architecture contributes to agent architectures over and
beyond \texttt{IDEA} in the following \kcomment{significant} ways:

@@

\item \emph{Partitioning} the global planning problem into multiple
  decision loops called {\em reactors}. Each reactor has its own scope
  both functionally (one for example focusing on the conversion of
  high level waypoints into low-level commands) and temporally as each
  reactor declares its planning look-ahead ($\pi$) and expected
  latency ($\lambda$) before being able to produce \kcomment{a} plan.
  \kcomment{These parameters depend} on where \kcomment{a reactor} is
  located in a dependency graph such as \kcomment{the one} shown in
  Fig. \ref{fig:agent}. A reactor closer to the hardware in the
  hierarchy, is expected to deliberate less and be highly responsive
  to observations from the world. This implies lower values of $\pi$
  and $\lambda$ (see Section \ref{sec:arch:trex} below). \kcomment{In
    the current design, $\pi$ and $\lambda$ are application dependent
    and fixed at design time.}

\item \emph{Coupling} a tighter integration between the planning
  process and world evolution \kcomment{with the use of the {\em same}
    \kcomment{plan} database}. By doing so, the planner is better
  informed in its deliberation despite a larger latency than the
  changing world.
\end{itemize}

Systematic partitioned inference is a major contribution of
\rxe. However that is not the focus of this chapter with details
covered in \cite{py10, rajan12}. Coupled interaction between planning
and the world on the other hand, has a substantial impact on how
planning can be integrated in such frameworks. In this chapter we show
how using the \eu planner can be coupled to the \rx framework by first
linking concepts in temporal planning with the framework. Note that in
the following treatise, we call the formal semantic framework of \rx
as the 'agent'. Reactors constitute an entity within this agent and
are standalone blocks of inference. The scope, detail and number of
reactors are application specific and outside the scope of this
chapter.

@@

The \rx architecture is structured around the notion of \eu {\em state
  variables} as the fundamental basis of interaction. Additionally it
comes with a well-defined ownership model associated with each such
state variable. In this framework each reactor is seen as a 'black
box' providing its own {\em internal} state variable which it is
responsible for in order to maintain its value at every instant of
time. Each such instance is represented by a {\em tick} with a fixed
duration; in our implementation of \rx on the Dorado AUV, we use a
tick of $1$ second. In order to identify its state a reactor can
subscribe to {\em external} state variables owned by other reactors,
for which it will receive updates when the owner reactor
\kcomment{makes changes} to its {\em internal} state \kcomment{due to
  deliberation or an exogenous state update}.

@@

As the execution frontier, $\tau$ advances, the \rx agent propagates
the current state information across all reactor state variables
through the {\em synchronization} process.  The sequencing of the
reactors synchronization is identified by traversing a reactor
dependency graph from the least dependent -- ones with no {\em
  external} state variables -- to the most, \ie which typically do not
have {\em internal} state variables. Conversely each reactor can
request a future change on one of its {\em external} state variables
which will be transformed as an {\em internal} goal to the owner
reactor(s).  Such \emph{goal dispatching} is expected to produce a
deliberation phase for the receiving reactor which can cyclically in
turn, produce a plan impacting its {\em external} state variables'
future that will propagate \kcomment{down the hierarchy} following the
same mechanism. Fig. \ref{fig:agent} shows an example of such a
dependency graph implemented as a directed acyclic graph.

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.45]{figs/token-disptch.pdf}
  \caption{\small \kcomment{Dispatching a token occurs only when the
      request start time overlaps a window. \texttt{Goal$_1$}
      therefore gets dispatched from \texttt{Reactor$_1$} to
      \texttt{Reactor$_2$} but not \texttt{Goal$_2$} to avoid early
      commitment. Tokens from \texttt{Reactor$_1$} are 'dispatched'
      for execution to \texttt{Reactor$_2$} only when necessary while
      meeting the conditions in Definition \ref{def:dispatch}. $\tau$
      signifies the execution frontier as time moves from left to
      right in the illustration.}}
  \label{fig:reactor-lookahead}
\end{figure}

@@

\item $\pi$ is the planning \emph{look-ahead} of the reactor. which is
  indicative of the number of ticks in the future this reactor is
  looking \kcomment{towards} deliberation.

@@

\begin{definition}
  The {\em execution latency} of a {\em state variable} $s$ which is
  internal to the reactor $r$ is the maximum expected delay before a
  goal on this state variable can be included in its plan and executed
  by the reactor $r$. It is defined recursively as: 
  \begin{equation*}
    \lambda_{exec}(s) = \lambda(r) + \max_{e \in exts(r)}
    \lambda_{exec}(e) 
  \end{equation*}
  Where  $exts(r)$ is the set of the state variables which are
  {\em external} to $r$. 
\end{definition}

For intertwined plan execution, the concept of execution latency is
critical at the agent level since it can impact the
planning/deliberation process on external state variables of a
reactor. For simplicity we abstract it out as being equivalent to the
reactor latency implying that we assume that all {\em external}
timelines of this reactor have $\lambda_{exec} = 0$. At the \rx agent
level this execution latency is used to define the {\em dispatch
  window} of the same state variable \kcomment{(see \cite{py10} for formal
proofs)}.

@@

This window is used by \rx to identify when to dispatch a request to a
{\em reactor} as a new goal. This happens only if the request start
time -- represented as the interval of possible start times --
intersects this window. Requests further in the future are buffered by
the agent until this time window advances for an overlap. This
provides an elegant mechanism to request a change from the {\em
  external} state variable of a reactor while avoiding flooding the
owner reactor of this state variable \kcomment{and avoiding early
  commitment.} The choice of both the latency and look-ahead of a
reactor can be delicate as one is dependent on the other. Latency is
dependent on exogenous paramaters such as the \kcomment{computer power
  of the host} and the number of reactors needing to deliberate at
once, among others. As the design stands, these variables should be
evaluated carefully by the system designer balancing agent
efficacy. Fig. \ref{fig:reactor-lookahead} illustrates an example
between two reactors where one token is dispatched when overlapping
the dispatch window while another is not.}

@@

\item The future objectives a reactor has on its \emph{internal} state
  variables will likely imply sub-objectives to be reported to the
  owner of its {\em external} state variables depending on their
  current values. Should the \textsf{Pilot} want to visit a specific
  \textsf{Waypoint} -- given its current position as provided by
  \textsf{Sensor\_Data}, it can identify a sequence of
  \textsf{Nav\_cmd} states that should eventually help it reach this
  location.

@@

Given the above example, \rx abstracts reactor interaction by
constraining them to state variables. Each reactor is therefore
agnostic where its {\em external} state information is managed as long
as this information is available and maintained for managing {\em
  internal} state information.  Such a formalism also constrains
reactor interaction to state information exchange which comes in two
forms as illustrated in Fig. \ref{fig:trex:flow}:

@@

\item {\em Observation on current state values}: This propagates
  \emph{up} in the reactor dependency graph \kcomment{during} the
  synchronization phase and occurs at the beginning of every
  tick. During this phase the agent provides updates to each reactor
  on their {\em external} state variable(s) to \kcomment{enable
    computation of} their {\em internal} state variable values for
  this same tick. By doing so we propagate a consistent view of world
  state throughout the reactors as time advances. \kcomment{Further,
    observations can be both sensor measurements as well as state
    information, passed up from other reactors. The strict rules of
    timeline ownership enable a clear policy for conflict resolution:
    observations dominate expectations.}

\item {\em Goal request on future state}: This information propagates
  \emph{down} in the reactor dependency graph. In order to satisfy its
  own {\em internal} objectives, a reactor may have a plan that relies
  on future values of one \kcomment{(or more)} of its {\em external}
  state. When such a plan is identified the agent ensures that this
  part of the plan is provided as a goal to the owner of the given
  {\em external} state variable(s).

@@

% The overall execution cycle of the agent is also abstracted out as a
\kcomment{Within \rx planning/deliberation is a continuous} cycle
between all reactors. At the agent interface level, each reactor
provides only two abstract calls that conceptualize execution:

@@

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example, one deliberation
  step corresponds \kcomment{to the resolution of} a single flaw
  (Section \ref{sec:europa:arch}) while ensuring that the resulting
  plan is not inconsistent. Backtracking is included in the step
  should it be \kcomment{necessary}.  The returned value by the
  \texttt{step} function is a boolean which indicates whether the
  reactor needs an extra step to deliberate or if this last step
  resulted in a complete plan within the reactor's
  look-ahead. \kcomment{Subsequently} the part of the plan that covers
  the reactor's {\em external} timelines is dispatched as a goal.

@@

Both these functions have different scope and temporal constraints.
The \texttt{synchronize} call is considered as atomic and focuses on
state inference for the current tick by identifying the current value
of {\em internal} state variables from the latest updates on {\em
  external} state variables. This process propagates up in the reactor
dependency graph at every tick to ensure that all reactors are aware
of current world state. \texttt{step} relates to a single decision
step of the planning process focused on the future evolution of state
variables; it can take multiple steps spanning multiple ticks for a
reactor to produce a plan that is both complete and valid. How the
reactor identifies that its plan is complete and valid is its
responsibility and depends on the way each reactor resolves its plan
at each {\em step}. The general idea is that the reactor is able to
find a valid strategy in order to satisfy as many of its {\em
  internal} goals as possible. For \eue, this means that the solver is
unable to identify any further flaws to be resolved in the
plan. \kcomment{Fig. \ref{fig:sync-cycle} illustrates the distinction
  between a tick, \texttt{step} and synchronization within the tick.}

When such a plan is identified, \kcomment{the reactors} {\em external}
timeline(s) are then dispatched to their \kcomment{respective reactor
  owners,} which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

% \label{sec:delib-exec-as}
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \subfloat[\small \kcomment{A view of synchronization and deliberation within a
  tick cycle.}]{\label{fig:sync-cycle}\includegraphics[width=0.35\columnwidth]{figs/sync-tick-cycle}} 
  \qquad
  \subfloat[\small A \rx reactor's execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.]{\label{fig:tick-exec}\includegraphics[width=0.45\columnwidth]{figs/tick-cycle}}
    \caption{\small \kcomment{Detailed and abstracted views of
        synchronization and deliberation within one and multiple tick
        cycles.}}
  \label{fig:sync-plan-cycle}
\end{figure}

@@

\subsubsection{Handling \kcomment{Reactor} Failure in \rx}
\label{sec:rx-reactor-failure}

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors \kcomment{and these timelines} are
\kcomment{then} considered \emph{rejectable} by the reactor that
manages them. As a result in such a situation, the failure to find a
plan for these goals, is \emph{not} considered a critical failure of
the reactor. It will only result \kcomment{in} these goals being
ignored until newer goals are received.

\begin{figure}[b]
  \centering \subfloat[\small When the \texttt{Sampling} reactor fails
  to synchronize, it is removed from the reactor dependency graph
  along with all its {\em internal} timelines while generating a
  \texttt{Failed} observation.]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.pdf}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize too leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent \kcomment{from} Fig. \ref{fig:agent}.}
\end{figure}

Take for example our AUV agent from Fig. \ref{fig:agent}, and assume
that the \texttt{Sampling} reactor failed to synchronize. In such a
case this reactor is destroyed and both \texttt{Sampling\_policy} and
\texttt{Relevant\_data} timelines generate the \texttt{Failed}
observation as illustrated in Fig. \ref{fig:fail0}.  The
\texttt{Survey controller} will receive this \texttt{Failed}
observation on its {\em external} \texttt{Sampling\_policy} timeline
which indicates that this state variable is no longer managed by its
owner reactor. This leads to two \emph{potential} outcomes based on
how the reactor is implemented:

@@

\item The \texttt{Survey controller} model \kcomment{could} integrate
  the \texttt{Failed} observation. For example it can identify that,
  subsequent use of the water sampler is no longer possible.  The \rx
  agent can continue to navigate since \texttt{Pilot} is still
  available. In such a case the agent structure remains as in
  Fig. \ref{fig:fail0} with no consequent impact.

@@

@@

\begin{figure}[!b]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    Fig. \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for \kcomment{and tokens in
        {\color{grey}grey} the result of deliberation}. Arrows
      indicate temporal  constraints \kcomment{with $==$ indicating 'equality' and $\geq$
        indicating 'greater than equal to' in timepoints}.}
  \label{fig:shop:exec}
\end{figure}

Consider that at the next tick $\tau' = \tau+1$ the reactor managing
agent location is able to change the location state variable to
\texttt{Going}. \rx informs the reactor of this observation which is
added to the partial plan as a fact -- starting exactly at $\tau'$ and
lasting for an unknown duration which is at least one tick. As this
new observation is integrated in the plan through synchronization, the
solver can identify the similarity between this observation and the
next token in the existing plan. Instead of initiating an exhaustive
search of all possible implications, the synchronization process
starts its execution with the assumption that these new observations
are consistent with the currently maintained plan. As a consequence,
at the end of synchronization the plan depicted in Fig.
\ref{fig:shop:exec1} is likely. This is a result of propagation of
information provided by the new observation in the existing plan while
restricting the start time of \texttt{Going} and propagating temporal
constraints, making \texttt{Go} to start {\em exactly} at $\tau'$. In
doing so, the \eu solver's focus on the execution frontier ($\tau'$ in
Fig. \ref{fig:shop:exec1}) takes advantage of the existing plan to
evaluate a solution that assumes that the observation is consistent
with what was decided by the reactor. \kcomment{In the event of
  inconsistency, the (plan) expectations do not match the reality;
  then partial plan failure within the reactor is initiated.} This is
directed not only by heuristic choices made in the solver but also
enforced by the heuristic order of the decision point resolution
choices for the {\em current state flaw} \kcomment{as} shown in
Proposition \ref{prop:csf:resolve}. As noted, the first two choices
are biased towards a conservative view with respect to the existing
plan.

\kcomment{\subsubsection{Dealing with partial plan failures}}
\label{sec:arch:partialfailures}

The \kcomment{solver's} heuristics first attempt to extend the
previous observation for this {\em internal} timeline to avoid early
commitment. If this fails \rx attempts to start the next token in the
plan to advance plan execution. Both choices attempt to resolve state
with the least amount of plan perturbation by constraining further the
start and end time of existing tokens generated by previous
deliberations steps. By doing so, in a nominal situation,
synchronization remains limited to restricting time points in the plan
and propagates their impact. Most of the computational effort of
creating new tokens in the plan and scheduling them onto timelines are
managed during deliberation steps. Consequently, most of the time
synchronization is limited to very few nodes in the \eu search space
often requiring no backtracking. Fig. \ref{fig:shop:exec} shows how
the existing plan can support the synchronization process. The
observation \texttt{Going} occurring at $\tau'$ is fully compatible
with the token already present in our plan. By updating the plan and
propagating equality constraints, the resolution of the {\em current
  state flaw} for the {\em Agent} timeline becomes obvious as a
solution, as the token \texttt{Go} from the plan presents itself.

@@

\begin{figure}[!b]
  \centering
  \subfloat[\small Conflicting Observation.]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[\small Relaxed Partial Plan.]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}
